---
title: "hw02-singhg53"
author: "Gurjot Singh"
date: "20/09/2018"
output: github_document
---

# Homework Assignment #2
#### By: Gurjot Singh (singhg53)

Hello! Today we will be exploring a the dataset known as `gapminder`

For those who have not yet downloaded the dataset you can do so by installing the package by the function:

`install.packages("gapminder")`

Now since you have downloaded the dataset lets load it using the `library` function

```{r}
library(gapminder)
```

Let's also install and load the tidyverse package

`install.packages("tidyverse")`

```{r}
library(tidyverse)
```


The cool thing about the tidyverse package is that includes different functions that could be used like `ggplot2` and `dplyr`

First let's check if the dataset is a vector:

```{r}
is.atomic(gapminder) || is.list(gapminder)
```

Now let's check what type of vector

```{r}
typeof(gapminder)
```

This shows us that the dataset is a list. But it is important to test if a dataset is a dataframe because a dataframe are lists as well.

```{r}
is.data.frame(gapminder)
```

This example shows good dimensionality of the dataset it is a vector-> list (1-dimension)-> data frame (2-dimension)


Let's just double check if the dataset is a matrix or array:

```{r}
is.matrix(gapminder)
is.array(gapminder)
```


```{r}
class(gapminder)
```




```{r}
nrow(gapminder)
```
