---
title: "hw02-singhg53"
author: "Gurjot Singh"
date: "20/09/2018"
output: github_document
---

# Homework Assignment #2
By: Gurjot Singh (singhg53)

Hello! Today we will be exploring a the dataset known as `gapminder`.

## Smell test the data

For those who have not yet downloaded the dataset you can do so by installing the package by the function:

`install.packages("gapminder")`

Now since you have downloaded the dataset lets load it using the `library` function:

```{r}
library(gapminder)
```

Let's also install and load the tidyverse package:

`install.packages("tidyverse")`

```{r}
library(tidyverse)
```

The cool thing about the tidyverse package is that includes different functions that could be used like `ggplot2` and `dplyr`

**Question: Is `gapminder` a data frame, a matrix, a vector, a list?**

First let's check what type of dataset is `gapminder` using the `typeof` function:

```{r}
typeof(gapminder)
```

`gapminder` is found to be a list.

This shows us that the dataset is a list. But it is important to test if a dataset is a dataframe because a dataframe are lists as well.

```{r}
is.data.frame(gapminder)
```

Since this is true we can establish that the dataset is indeed a data frame.

A cool way to check if the dataset is a martix or an array is using the functions below.

```{r}
is.matrix(gapminder)
is.array(gapminder)
```

Both of them come out false, as expected.

**Question: What is its class?**

The class of the dataset can be found by using the `class` function

```{r}
class(gapminder)
```

This is another way we could find that `gapminder` is a data frame.

**Question: How many variables/columns?**

The number of columns can be found using the `ncol` function:

```{r}
ncol(gapminder)
```

This shows us that there are six columns in the dataset.

**Question: How many rows/observations?**

Let's check how many rows are in the data set using the `nrow` function:

```{r}
nrow(gapminder)
```

We can conclude there are 1704 rows in this dataset.

**Question: Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?**

We can use the `dim` function to tell us the size or the exent of the data

```{r}
dim(gapminder)
```

The `dim` function could give us the number of rows and columns using one function, where the number of rows come first and the number of columns comes second.

**Question: What data type is each variable?**

The structure of the dataset can also tell us about the extent or size of the data and it could tell us what data type is each variable

```{r}
str(gapminder)
```

The data type of each variable has been summarized in the table below:

|    **Variable**      |    **Data Type** |
|----------------------|------------------|
|       country        |     Factor       |
|       continent      |     Factor       |
|        year          |     Integer      |
|       lifeExp        |     Number       |
|        pop           |    Integer       |
|       gdpPercap      |     Number       |

# Explore Individual Variables

Question: Pick **at least** one categorical variable and at least one quantitative
variable to explore.

  - What are possible values (or range, whichever is appropriate) of
    each variable?
  - What values are typical? What’s the spread? What’s the distribution?
    Etc., tailored to the variable at hand.
  - Feel free to use summary stats, tables, figures. We’re NOT expecting
    high production value (yet).

Let's first explore the continent variable, which is categorical:

```{r}
summary(gapminder$continent)
```

Using the `summary` function we could see that the continent dataset has five possible variables: Africa, Americas, Asia, Europe, and Oceania. In addition, the `summary` function it tells us information about the values of each variable. Africa has 624 observations (rows of information) versus the Americas which have 300 observations.

Now let's look at a numerical variable such as lifeExp

```{r}
summary(gapminder$lifeExp)
```

Again, the `summary` function is very powerful because it can provide us with a great deal of information. Using this function we have found the spread of the data giving the maximum (82.60), the minimum (23.60), the median (60.71), and the mean (59.47).

Let's do another numerical variable for fun to go above and beyond, such as, gdpPercap

```{r}
summary(gapminder$gdpPercap)
```

Using this function we have found the spread of the data giving the maximum (113523.1), the minimum (241.2), the median (3531.8), and the mean (7215.3).

## Explore various plot types

Question: Make a few plots, probably of the same variable you chose to
characterize numerically. You can use the plot types we went over in
class (cm006) to get an idea of what you’d like to make. Try to explore
more than one plot type. Just as an example of what I mean:

  - A scatterplot of two quantitative variables.
  - A plot of one quantitative variable. Maybe a histogram or
    densityplot or frequency polygon.
  - A plot of one quantitative variable and one categorical. Maybe
    boxplots for several continents or countries.

You don’t have to use all the data in every plot\! It’s fine to filter
down to one country or small handful of countries.

Use `filter()`, `select()` and `%>%`

Use `filter()` to create data subsets that you want to plot.

Practice piping together `filter()` and `select()`. Possibly even piping
into `ggplot()`.

**Question: A scatterplot of two quantitative variables.**

First let's make a scatterplot of two quantitative variables. I shall use lifeExp and gdpPercap.

```{r}
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  scale_x_log10() +
  geom_point(colour = "blue", alpha=0.2)
```

I have changed the y-axis to make it a logarithmic scale for better visualization and added some colour.

Let's take this a step further and use the `pipe` function and the `filter` function.

```{r}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot(aes(gdpPercap, lifeExp)) +
  geom_point(colour = "blue", fill = "blue")
```

**Question: A plot of one quantitative variable. Maybe a histogram or densityplot or frequency polygon.**

Let's use a different country, for example, Mexico, and plot a histogram of the lifeExp:

```{r}
gapminder %>%
  filter(country == "Mexico") %>%
  ggplot(aes(lifeExp)) +
  geom_histogram(colour = "red", fill = "orange", bins = 10)
```

Let's go above and beyond and create a density plot for a different country, such as, Canada and observing lifeExp.

```{r}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot(aes(lifeExp)) +
  geom_density(colour = "red")
```

**A plot of one quantitative variable and one categorical. Maybe boxplots for several continents or countries.**

In this plot we filtered out years greater than 2000 and plotted the continent with the lifeExp using a violin boxplot. This type of boxplot gives great information regarding the overall distribution of the data.

```{r}
gapminder %>% 
  filter(year > 2000) %>% 
  ggplot(aes(continent, lifeExp)) +
  geom_boxplot() +
  geom_violin(width = 1, fill = "orange", alpha=0.5)
```

**A little something extra**

Now let's do something that hasn't been done in class yet. We are going to create a 2d density plot using variables of lifeExp and gdpPercap!

```{r}
ggplot(gapminder, aes(lifeExp, gdpPercap)) +
  geom_point(alpha = 0.5) + 
  geom_density_2d() + 
  scale_y_log10()
```

## Resources

[Here](http://genomicsclass.github.io/book/pages/dplyr_tutorial.html) is a site I used to help create the plots. It is a great resource!

## But I want to do more\!

*For people who want to take things further.*

Evaluate this code and describe the result. Presumably the analyst’s
intent was to get the data for Rwanda and Afghanistan. Did they succeed?
Why or why not? If not, what is the correct way to do this?

    filter(gapminder, country == c("Rwanda", "Afghanistan"))

The above would be the incorrect way of filtering out information for the countries. Using the technique above you loose a great deal of information, becuase the information is known to be "recycled".

To circumvent this problem we could do the following operation:

```{r}
filter(gapminder, country == "Rwanda" | country == "Afghanistan")
```

I hope you enjoyed going through my homework assignment!