---
title: "hw02-singhg53"
author: "Gurjot Singh"
date: "20/09/2018"
output: github_document
---

# Homework Assignment #2
#### By: Gurjot Singh (singhg53)

## Smell test the data

Hello! Today we will be exploring a the dataset known as `gapminder`

For those who have not yet downloaded the dataset you can do so by installing the package by the function:

`install.packages("gapminder")`

Now since you have downloaded the dataset lets load it using the `library` function

```{r}
library(gapminder)
```

Let's also install and load the tidyverse package

`install.packages("tidyverse")`

```{r}
library(tidyverse)
```


The cool thing about the tidyverse package is that includes different functions that could be used like `ggplot2` and `dplyr`

First let's check if the dataset is a vector:

```{r}
is.atomic(gapminder) || is.list(gapminder)
```

Now let's check what type of vector

```{r}
typeof(gapminder)
```

This shows us that the dataset is a list. But it is important to test if a dataset is a dataframe because a dataframe are lists as well.

```{r}
is.data.frame(gapminder)
```

This example shows good dimensionality of the dataset it is a vector-> list (1-dimension)-> data frame (2-dimension)


Let's just double check if the dataset is a matrix or array:

```{r}
is.matrix(gapminder)
is.array(gapminder)
```

The class of the dataset can be found by using the `class` function

```{r}
class(gapminder)
```

The columns can be found using the `ncol` function

```{r}
ncol(gapminder)
```

Let's check how many rows are in the data set using the `nrow` function

```{r}
nrow(gapminder)
```

We can use the `dim` function to tell us the size or the exent of the data

```{r}
dim(gapminder)
```

The structure of the dataset can also tell us about the extent or size of the data and it could tell us what data type is each variable

```{r}
str(gapminder)
```


|    **Variable**      |    **Data Type** |
|----------------------|------------------|
|       country        |     Factor       |
|       contient       |     Factor       |
|        Year          |     Integer      |
|       lifeExp        |     Number       |
|        pop           |    Integer       |
|       gdpPercap      |     Number       |

# Explore Individual Variables

Let's first explore the continent variable, which is categorical

```{r}
summary(gapminder$continent)
```

Now let's look at a numerical variable such as lifeExp

```{r}
summary(gapminder$lifeExp)
```

Let's do another cateogrical variable for fun, such as, continent

```{r}
summary(gapminder$continent)
```

