---
title: "hw02-singhg53"
author: "Gurjot Singh"
date: "20/09/2018"
output: github_document
---

# Homework Assignment #2
By: Gurjot Singh (singhg53)

Hello! Today we will be exploring a the dataset known as `gapminder`.

## Smell test the data

For those who have not yet downloaded the dataset you can do so by installing the package by the function:

`install.packages("gapminder")`

Now since you have downloaded the dataset lets load it using the `library` function:

```{r}
library(gapminder)
```

Let's also install and load the tidyverse package:

`install.packages("tidyverse")`

```{r}
library(tidyverse)
```

The cool thing about the tidyverse package is that includes different functions that could be used like `ggplot2` and `dplyr`

**Question: Is `gapminder` a data frame, a matrix, a vector, a list?**

First let's check what type of dataset is `gapminder` using the `typeof` function:

```{r}
typeof(gapminder)
```

`gapminder` is found to be a list.

This shows us that the dataset is a list. But it is important to test if a dataset is a dataframe because a dataframe are lists as well.

```{r}
is.data.frame(gapminder)
```

Since this is true we can establish that the dataset is indeed a data frame.

A cool way to check if the dataset is a martix or an array is using the functions below.

```{r}
is.matrix(gapminder)
is.array(gapminder)
```

Both of them come out false, as expected.

**Question: What is its class?**

The class of the dataset can be found by using the `class` function

```{r}
class(gapminder)
```

This is another way we could find that `gapminder` is a data frame.

**Question: How many variables/columns?**

The number of columns can be found using the `ncol` function:

```{r}
ncol(gapminder)
```

This shows us that there are six columns in the dataset.

**Question: How many rows/observations?**

Let's check how many rows are in the data set using the `nrow` function:

```{r}
nrow(gapminder)
```

We can conclude there are 1704 rows in this dataset.

**Question: Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?**

We can use the `dim` function to tell us the size or the exent of the data

```{r}
dim(gapminder)
```

The `dim` function could give us the number of rows and columns using one function, where the number of rows come first and the number of columns comes second.

**Question: What data type is each variable?**

The structure of the dataset can also tell us about the extent or size of the data and it could tell us what data type is each variable

```{r}
str(gapminder)
```

The data type of each variable has been summarized in the table below:

|    **Variable**      |    **Data Type** |
|----------------------|------------------|
|       country        |     Factor       |
|       contient       |     Factor       |
|        Year          |     Integer      |
|       lifeExp        |     Number       |
|        pop           |    Integer       |
|       gdpPercap      |     Number       |

# Explore Individual Variables

Question: Pick **at least** one categorical variable and at least one quantitative
variable to explore.

  - What are possible values (or range, whichever is appropriate) of
    each variable?
  - What values are typical? What’s the spread? What’s the distribution?
    Etc., tailored to the variable at hand.
  - Feel free to use summary stats, tables, figures. We’re NOT expecting
    high production value (yet).

Let's first explore the continent variable, which is categorical

```{r}
summary(gapminder$continent)
```

Using the `summary` function we could see that the continent dataset has five possible variables: Africa, Americas, Asia, Europe, and Oceania. In addition, the `summary` function it tells us information about the values of each variable. Africa has 624 observations (rows of information) versus Americas has 300 observations.

Now let's look at a numerical variable such as lifeExp

```{r}
summary(gapminder$lifeExp)
```

Again, the `summary` function is very powerful because it can provide us with a great deal of information. Using this function we have found the spread of the data



Let's do another cateogrical variable for fun, such as, continent

```{r}
summary(gapminder$continent)
```







Vectorization, recycling

look at the years

```{r}
filter(gapminder, country == "Rwanda" & country == "Afghanistan")
```








```{r}
filter(gapminder, country == c("Rwanda"))
```


